File-based localization of numerical perturbations in data analysis pipelines

### 2020-02-24 paper.tex
Working on the method

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip\xspace}
\newcommand{\tristan}[1]{\color{blue}\textbf{From Tristan:}#1\color{black}}
\newcommand{\english}[1]{\uwave{#1}}

@@

Spot identifies processes that create differences from a process graph of the pipeline execution.
This tool compares file dependencies through all the processes based on the result files in two conditions.
In this paper, different conditions refer to various operating systems.
This tool is applicable on the complex pipelines in which tracing a large number of processes is difficult.

Figure~\ref{fig:overview-tool} shows an overview of our method. For each tested
condition, Condition 1 and Condition 2, the pipeline is containerized with
Docker, and its interface is described with
Boutiques~\cite{glatard2017boutiques}. 
\reprozip is added to Condition 1 for provenance tracking. 
This captures process dependencies through system call interceptions, 
then Spot tool creates related process trees (Figure~\ref{fig:overview-tool} - (1)).
Different subjects produce different process trees. 
Hierachical clustering used to group subjects into different types (Figure~\ref{fig:overview-tool} - (2)).
The rest of the procedure is done separately on each subject type.
Our tool continues by
restoring temporary files and ``multi-write" files, i.e., files written
by more than one process (Figure~\ref{fig:overview-tool} - (3)). 
To do so, the tool creates a list of processes 
that introduce intermediary files and replaces their corresponding executables with a wrapper script. 
The tool re-executes the pipeline in a modified Docker image in Condition 1.
This copies files in the list to a read-only directory.
In the next step, the
processes are labeled from the differences found in their outputs in the
two tested conditions (Figure~\ref{fig:overview-tool} - (4)). 
The tool builds the Docker image for Condition 2 by modifying all
the processes in the pipeline. Then, the modified Docker
image is executed in Condition 2 and each process is labeled after
running by comparing the results obtained in Condition 1
and output files of the process in Condition 2. We label processes in 2
categories depending on whether they
create differences (represented in red) or not (green). 
The tool replaces output file with the same file obtained in Condition 1 
if the process creates differences.
Process labelling
is done through hooks which are defined during the Docker image modifying
procedure in Condition 2.
Spot tool is available at https://github.com/big-data-lab-team/repro-tools.

@@

We use \reprozip 
to capture: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call and stores it in a SQLite database.

Our tool reconstructs a \emph{process tree} starting from the first process
created by the pipeline and traversing processes through \texttt{clone()}
and \texttt{fork()} dependencies. Our tool also creates a process \emph{graph} 
from this tree
by adding edges corresponding to file dependencies between processes. A
file dependency is defined between processes A and B if a file written by A
is read by B. Figure~\ref{fig:simple_script} shows an example of a process
tree and graph constructed from the example pipeline in
Listing~\ref{listing:sample-script}.

Different subjects may lead to different process trees, due to
variations that may be coming from differences in data types or
cardinality. For instance, some images may have been acquired multiple
times in some subjects. Some of these differences can be neglected, for
instance, when a data decompression step is present at the beginning of the
execution for some subjects only, and other ones cannot, when different
processing paths are used in different subjects.
Accordingly, we cluster subjects into different types.

\begin{listing}
\begin{minted}[frame=single,
  framesep=3mm,
  linenos=false,
  xleftmargin=0pt,
  tabsize=4]{bash}
#!/bin/bash

# This script extracts the brain from an input image,
# measures the volume of the extracted brain,
# and creates a binarized brain mask.

if [ $# != 1 ]
then
    echo "usage: $0 <inputimage.nii.gz>"
    exit 1
fi

# Parse argument, set file names
input_image=$1
bet_output=$(basename "${input_image}" \
                       .nii.gz)_brain.nii.gz
bet_output_binarized=$(basename "${input_image}" \
                       .nii.gz)_brain_bin.nii.gz

# Run FSL bet, put result in ${bet_output}
bet "${input_image}" "${bet_output}" > bet_temp.out
echo "Voxels / Volume in brain mask:"
# Run FSL stats on ${bet_output}
fslstats "${bet_output}" -V
# Run FSL maths on ${bet_output}
fslmaths "${bet_output}" "${bet_output_binarized}"
# Remove temporary file
\rm bet_temp.out
\end{minted}
  \caption{Example pipeline}
  \label{listing:sample-script}
\end{listing}

@@

We cluster subjects into different types to specify how many different 
process trees are generated. 
Nodes in the process trees are labeled by the process name.
The tree edit distance~\cite{zhang1989simple} is used as a similarity measure for 
representing the distance between labeled trees. 
Tree edit distance  is defined  
as the minimum number of edit operations which is needed to transform one tree into 
the other. Three edit operations are considered: node label modification, 
node removal, and node insertion. Each operation has an associated cost of 1.
With this distance, we cluster subjects using agglomerative hierarchical
clustering as implemented in SciPy~\cite{oliphant2007scipy}.
As a result, the distance between
any two subjects in a cluster is lower than a threshold parameter.

We will show that different types of subjects certainly produce different results, 
while subjects in the same type can be either reproducible or not.

>>>>

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip\xspace}
\newcommand{\tristan}[1]{\color{blue}\textbf{From Tristan:}#1\color{black}}
\newcommand{\english}[1]{\uwave{#1}}
\newcommand{\toolname}[0]{Spot\xspace}

@@

Figure~\ref{fig:overview-tool} shows an overview of our method. The output
of 2 computational conditions is compared through 3 main steps. At step 1,
we build a low-level provenance graph for the processing of each subject in
Condition 1, where nodes are OS processes created by the application and
edges are file dependencies. Such provenance graphs reconstruct a model of
the application workflow, regardless of the original programming language
and dependencies. They may, however, slightly differ across the processed
subjects since different execution paths might be followed due to
variations in the data. To address this issue, we cluster subjects based on
their provenance graph (step 2), resulting in a reduced set of $k$
representative provenance graphs. These $k$ graphs and their corresponding
subjects /TODO: ???/ are then used for process labeling (step 3), where
Condition 2 is executed step by step, comparing the files produced by each
process to their counterparts in Condition 1, making sure that temporary
files and files written by more than a single process are correctly saved.

@@

We use the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call and stores it in a SQLite database.

\toolname reconstructs a \emph{process tree} starting from the first process
created by the pipeline and traversing processes through \texttt{clone()}
and \texttt{fork()} dependencies. It also creates a process \emph{graph} 
from this tree by adding edges corresponding to file dependencies between processes. A
file dependency is defined between processes A and B if a file written by A
is read by B. Figure~\ref{fig:simple_script} shows an example of a process
tree and graph constructed from the example pipeline in
Listing~\ref{listing:sample-script}.

\begin{listing}
  \inputminted{bash}{"bin/example.sh"}
  \caption{Example pipeline}
  \label{listing:sample-script}
\end{listing}

@@

Different subjects may lead to different provenance graphs, due to
variations that may be coming from differences in data types or
cardinality. For instance, some images may have been acquired multiple
times in some subjects, for noise-reduction purposes. Some of these
differences can be neglected, for instance, when a data decompression step
is present at the beginning of the execution for some subjects only, but
other differences cannot be ignored, for instance when complete different
processing paths are used in different subjects.

To reduce the number of provenance graphs to be examined, we cluster
process trees using agglomerative hierarchical clustering, as implemented
in SciPy~\cite{oliphant2007scipy}. We use the tree edit
distance~\cite{zhang1989simple} between process trees, as implemented in
the zss Python package. This distance is defined as the minimum number of
edit operations to transform one tree into the other. Three edit operations
are considered: node label modification, node removal, and node insertion.
Each operation has an associated cost of 1. 
//If threshold is 0, why do we even need the edit distance?//
